Approximation as Differentiation

### 2025-05-07 main/formal-structures.tex
Biproducts and semi-additive categories.
Galois connections.

>>>>

\subsection{Category of lattices and Galois connections}

First we introduce the category of lattices and Galois connections, which will serve as our base category for
interpreting forward and backward Galois slicing. Galois connections are pairs of monotone functions between
posets, where each is the closest approximation to an inverse of the other: one from above, one from below.

\begin{definition}[Galois connection]
Suppose $X$ and $Y$ are posets. A \emph{Galois connection} $f \adj g: X \to Y$ is a pair of monotone functions
$f: X \to Y$ and $g: Y \to X$ satisfying $y \leq f(x) \iff g(y) \leq x$ for any $x \in X$ and $y \in Y$.
\end{definition}

\noindent The $\adj$ notation is justified because a Galois connection $f \adj g: X \to Y$ can also be seen an
adjunction between poset-as-categories, with monotone $f$ and $g$ interpreted as functors; $f$ is usually
referred to as the \emph{upper} (right) adjoint and $g$ as the \emph{lower} (left) adjoint. Galois connections
compose component-wise, with $\id_X \adj \id_X: X \to X$ as the unit for composition. Thus we can define the
following category.
